---
title: "assignment"
author: "Guillermo Turiel"
date: "2023-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(GenomicRanges)
  library(epiwraps)
  library(rtracklayer)
  library(ggplot2)
})
proxy_url <- "http://proxy.ethz.ch:3128";
Sys.setenv(http_proxy = proxy_url, https_proxy = proxy_url, ftp_proxy = proxy_url)
setwd("C:/Users/guruiz/Desktop/ethEpigen2023/week05")
```
## Loading files from mESC
```{r}
H3K27me3_peaks <- rtracklayer::import("C:/Users/guruiz/Desktop/ethEpigen2023/week04/H3K27me3.bed.gz", format="narrowPeak")
H3K4me3_peaks <- rtracklayer::import("C:/Users/guruiz/Desktop/ethEpigen2023/week04/H3K4me3.bed.gz", format="narrowPeak")

```
## Subsetting bivalent domains in mESC
```{r}
bivalent_mesc <- H3K4me3_peaks[overlapsAny(H3K4me3_peaks, H3K27me3_peaks)]
print(paste(length(bivalent_mesc), "peaks from H3K4me3 are bilavent with H3k27me3 peaks in mESC"))
```


## Downloading H3K27me3 and H3K4me3 peaks from hepatocytes
```{r}
download.file("https://www.encodeproject.org/files/ENCFF251NKN/@@download/ENCFF251NKN.bed.gz", dest="H3K27me3_hep.bed.gz")
H3K27me3_hep <- rtracklayer::import("H3K27me3_hep.bed.gz", format="narrowPeak")

download.file("https://www.encodeproject.org/files/ENCFF692VTG/@@download/ENCFF692VTG.bed.gz", dest="H3K4me3_hep.bed.gz")
H3K4me3_hep <- rtracklayer::import("H3K4me3_hep.bed.gz", format="narrowPeak")
```

## Checking overlaps between bivalent marks in mESC and the marks in hepatocytes
```{r}
peaks = list(bivalent=bivalent_mesc, H3K27me3_hep=H3K27me3_hep, H3K4me3_hep=H3K4me3_hep)
regionUpset(peaks, reference = peaks$bivalent, nsets=length(peaks))
```

From the upset plot we get the following conclusions:
- 345 of the bivalent peaks in mESC dont overlap any mark in hepatocytes.
- 1149 bivalent peaks overlap only with H3K27me3 peaks in hepatocytes.
- 957 bivalent peaks overlap only with H3K4me3 peaks in hepatocytes.
- 1143 bivalent peaks overlap with both H3K27me3 and H3K4me3 peaks in hepatocytes.

