---
title: "assignment"
author: "Guillermo Turiel"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(GenomicRanges)
  library(epiwraps)
  library(ggplot2)
  library(rGREAT)
})
proxy_url <- "http://proxy.ethz.ch:3128";
Sys.setenv(http_proxy = proxy_url, https_proxy = proxy_url, ftp_proxy = proxy_url)
setwd("C:/Users/guruiz/Desktop/ethEpigen2023/week10")
```


## Download the data

```{r, eval=FALSE}
options(timeout = 6000)
download.file("https://ethz-ins.org/content/w10.assignment.zip", "w10.assignment.zip")
unzip("w10.assignment.zip")
```

## Prepare the regions and the tracks

```{r}
tracks <- list.files(pattern="bw$")
peaks <- list.files(pattern="bed$")
# we first import the peaks
peaks <- lapply(peaks, rtracklayer::import.bed)
# we'll focus on the high-quality peaks
peaks <- lapply(peaks, FUN=function(x) x[x$score>800])
# we get the union of non-redundant regions
regions <- reduce(unlist(GRangesList(peaks)))
```

## Plot the signals

```{r}
ml <- signal2Matrix(tracks, regions, extend=2000)

plotEnrichedHeatmaps(ml)
```

## Clustering

```{r}
k=4
set.seed(123)
cls <- clusterSignalMatrices(ml, k=k)
mycolors <- scales::hue_pal()(k)
names(mycolors) = 1:k
plotEnrichedHeatmaps(ml, row_split=cls, mean_color=mycolors)
d <- meltSignals(ml, splitBy=cls)
ggplot(d, aes(position, mean, colour=sample)) + geom_line() + facet_wrap(~split)
```

Interpretation of the clusters:

- Cluster 1: a set of regions mainly bound by Creb1, although a little signal is also detected for Creb3L1.

- Cluster 2: a set of regions mainly bound by all 3 factors.

- Cluster 3: a set of regions that are only bound by Creb3 and not any other factor.

- Cluster 4: a set of regions mainly bound by Creb1 and Creb3L1.

From this results we can conclude that Creb1 and Creb3L1 often bind to the same regions (clusters 1 and 4) while Creb3 mainly binds to other regions in an independent manner (cluster 3). However, a small set of regions is also bound by all 3 factors (cluster 2). 

## Enrichment analysis

```{r}
# we first split the regions by cluster:
split_regions <- split(regions, cls)
# we send the data to the GREAT server to run an enrichment analysis
#We focus on cluster 3
job <- submitGreatJob(gr=split_regions[["3"]], bg=regions, species="hg38")
res <- getEnrichmentTables(job)
```

We plot the top Molecular Functions:

```{r, fig.width=9, fig.height=6}
bp <- res$`GO Molecular Function`
head(bp)
ggplot(head(bp,15), aes(Hyper_Fold_Enrichment, reorder(name, Hyper_Adjp_BH), size=Hyper_Foreground_Region_Hits, color=-log10(Hyper_Adjp_BH))) + geom_point() + scale_color_viridis_c()
```


The enrichment analysis on cluster 3 (regions only bound by Creb3) revealed that these regions are related to genes involved in receptor binding and activity. These receptors are mainly related to cytokines such as TNF.
